Layered Formal Verification of a TCP Stack

### 2021-08-26 SecDev21/paper.tex
fix biblio
Corrections for Fig. 3 and Algorithm 1

To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}. It computes the set of possible final states after the completion of an event by calling repeatedly the function \spark{TCP_Process_One_Segment}. Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}. This ghost function is proved automatically without even requiring a loop invariant, thanks to loop unrolling in GNATprove.

@@

// Check the expected postcondition
klee_assert( (*@\label{code:kleedriver:assert}@*)
    (oldSock.state == TCP_STATE_ESTABLISHED) ?
        sock.state == TCP_STATE_ESTABLISHED ||
        sock.state == TCP_STATE_CLOSE_WAIT ||
        sock.state == TCP_STATE_CLOSED :
    (oldSock.state == ...) ? ... )
)
\end{lstlisting}
\caption{Driver for the verification of \lstinline[language=C]{tcpProcessSegment}
with KLEE}
\label{code:kleedriver}
\end{figure}

@@

These results depend upon other layers in the C programming language, which were not formally verified to the same level. Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors. The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer. This processing part can be a source of errors and bugs as explained in~\cite{Reiher2019RecordFluxFM}. As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification. This would complete our current verification of the CycloneTCP stack towards the lower layers.

>>>>

To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}. It computes the set of possible final states after the completion of an event by calling repeatedly the function \spark{TCP_Process_One_Segment}. It is enough to perform only three iterations in the loop of this algorithm because the maximum path between two states that does not require a user interaction has a size of three. Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}. This ghost function is proved automatically without even requiring a loop invariant, thanks to loop unrolling in GNATprove.

@@

// Check the expected postcondition
klee_assert( (*@\label{code:kleedriver:assert}@*)
    (oldSock.state == TCP_STATE_ESTABLISHED) ?
        sock->state == TCP_STATE_ESTABLISHED ||
        sock->state == TCP_STATE_CLOSE_WAIT ||
        sock->state == TCP_STATE_CLOSED :
    (oldSock.state == ...) ? ... )
)
\end{lstlisting}
\caption{Driver for the verification of \lstinline[language=C]{tcpProcessSegment}
with KLEE}
\label{code:kleedriver}
\end{figure}

@@

These results depend upon other layers written in the C programming language, which were not formally verified to the same level. Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors. The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer. This processing part can be a source of errors and bugs as explained in~\cite{Reiher2019RecordFluxFM}. As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification. This would complete our current verification of the CycloneTCP stack towards the lower layers.
